### 2015-07-09 text/ms.Rmd
More edits to paper

\noindent
Sean C. Anderson^1\*^, 
Andrew B. Cooper^1^
Olaf P. Jensen^4^, 
C\'{o}il\'{i}n Minto^6^, 
James T. Thorson^11^,
Jessica C. Walsh^1^,
Jamie Afflerbach^2^, 
Mark Dickey-Collas^3^, 
Kristin M. Kleisner^5^, 
Catherine Longo^2^, 
Giacomo Chato Osio^7^, 
Dan Ovando^8^, 
Andrew A. Rosenberg^9^, 
Elizabeth R. Selig^10^, 
*(order to be determined and superscripts out of order)*

@@

Ecological resource managers are often faced with reconciling multiple potentially
conflicting estimates of population status and trajectory.  One solution is to
average these predictions in an ensemble, but this approach ignores covariance
between models and may not optimally reduce bias and improve accuracy.
Superensemble models, commonly used for climate and weather
forecasting, may provide a superior solution. Superensembles use the predictions
from multiple models as covariates in a new statistical model. Here we evaluate
the potential for ensemble and superensemble models ('ensemble methods') to improve estimates of fish population
status and trajectory. We fit four data-limited models of population and exploitation
dynamics to marine fish stocks and combine their estimates of the mean and
slope of biomass at maximum sustainable yield ($B/B_\mathrm{MSY}$) with an
ensemble average and three superensembles: a linear model, a random forest, and
a boosted regression tree. 
We selected data-limited models for their applicability across the majority of stocks world wide.
We build our models on a simulated dataset of 5760
fish stocks and tested them with cross-validation and a global database of 
`r ram_stocks_n` stock assessments that have also been assessed with more data-intensive stock assessment models.
We find that ensemble methods substantially improve estimates of population status and trajectory.
Machine-learning superensembles performed the best for population status: accuracy improved
`r mean_sim$mach_vs_ind_mare_fold`%, correlation between predicted and true
status improved from 
`r mean_sim$ind_corr_range` to `r mean_sim$mach_corr_range`, and bias (median
proportional error) declined from 
`r mean_sim$ind_mre_range` to `r mean_sim$mach_mre_range`. 
We found similar improvements when predicting the trajectory of status. When the
simulation-trained superensembles were applied to catch data for global fish stocks, the best
superensembles improved all dimensions of performance over the next-best
individual model. Ensemble methods have the potential to
improve estimates of population status and trajectory across many taxa beyond
fish. However, they must be carefully tested, formed from a diverse set of
accurate models, and built on a dataset sufficiently representative of the
populations to which they are applied.

@@

Population status and trajectory are two of the most fundamental values to quantify in ecological resource management [e.g. @mace2008; @iucn2015]. These properties are key to everything from global assessments of extinction risk [e.g. @stuart2004; @schipper2008; @juan-jorda2011] to setting hunting (REF) and fishing limits [e.g. @worm2009] (REFs). However, ecological resource managers are often faced with reconciling multiple uncertain and potentially conflicting estimates of status and trajectory [e.g. @branch2011] (REFs). Some models may suggest a population is at risk and declining in abundance while others may suggest it is not at risk and stable.

@@

In fisheries science, the gold standard for determining status and trajectory
of exploited fish populations are stock assessments that
incorporate all available data (e.g., catches, size and age distributions,
surveys, and tagging information) to quantify values such as 
the biomass that can produce maximum sustainable yield ($B/B_\mathrm{MSY}$).
However, these kinds of data are not available for the majority of stocks
[@fao2014]. Therefore, a number of models have been proposed to assess
$B/B_\mathrm{MSY}$ based on the limited data available for the majority of
stocks: catches and a basic understanding of population productivity [e.g.
@vasconcellos2005; @martell2013]. Recently, @rosenberg2014 investigated the
performance of four data-limited models through a large-scale simulation
experiment. The four models frequently disagreed about population status and
any one model had relatively poor predictive ability on average (e.g. Fig.
\ref{motivate}).

@@

* *mPRM* (modified panel regression model) is a modified version of the
  panel-regression model from @costello2012. Unlike the other
  models, mPRM is empirical and not mechanistic---it uses the RAM Legacy Stock
  Assessment database to fit a regression model to a series of characteristics
  of the catch time series and stock with stock-assessed $B/B_\mathrm{MSY}$ as
  the response. The model used in this paper is modified from the original in
  that it condenses the life-history categories into three categories to match
  the simulated dataset, removes the maximum catch predictor since the absolute
  catch in the simulated dataset is arbitrary, and does not implement the bias
  correction needed in @costello2012 for deriving aggregate
  estimates of mean status across multiple stocks.

@@

Predictive performance can be evaluated with metrics that represent a variety
of modelling goals. For continuous response variables such as the mean and
slope of population status, performance metrics often measure some form of
bias, precision, accuracy (a combination of bias and precision), or the ability
to correctly rank or correlate across populations [e.g. @walther2005]. Here,
we measure proportional error, defined as
$(\hat{b} - b)/b$, where $\hat{b}$ and $b$ represent estimated and 'true'
(or stock-assessed) mean or slope of $B/B_\mathrm{MSY}$. We calculated median proportional
error to measure bias, median absolute proportional error to measure accuracy,
and Spearman's rank-order correlation to measure the ability to correctly rank
populations.

@@

<!--TODO: check labels on Figure S4 – repeat of (d) and missing (f)-->
The superensemble models were able to improve the predictions of status by
exploiting the best properties of individual models, the covariance between
individual models, and interactions with other covariates. For example, SSCOM
had strong predictive ability when it predicted low $B/B_\mathrm{MSY}$ (Fig.
\ref{hexagon}c, Fig. \ref{partial-sim}c) and CMSY predictions were
approximately linearly related to $B/B_\mathrm{MSY}$ within the low and high prediction
clusters (Fig. \ref{partial-sim}). Superensembles also exploited the covariance
between individual model predictions. For instance, both the linear model and
GBM ensemble suggest that if mPRM and SSCOM predict high status, the true
status also tends to be high (Figs \ref{lm-coefs}, \ref{partial-2d-sim}l). The
addition of spectral density covariates helped the superensemble models correctly
predict higher status values (Fig. \ref{hexagon}g, h) by exploiting
interactions between the spectral density predictors and the individual model
predictions (Fig. \ref{partial-2d-sim}). Although the additional covariates
improved ensemble fit, performance of the ensembles was only marginally
degraded by removing these covariates (Fig. \ref{hexagon} vs. Fig.
\ref{hexagon-sim-basic}).

@@

A strength of superensembles is that they can be tailored to predict specific response
variables. For example, we built separate superensemble models of mean
$B/B_\mathrm{MSY}$ and the slope of $B/B_\mathrm{MSY}$. The same set of model
weights or non-linear relationships need not hold across different response
variables. For example, SSCOM contributed little to the GBM superensemble estimate
of status at higher levels of predicted $B/B_\mathrm{MSY}$, but contributed
strongly to estimates of slope (TODO FIG). Formally, fitting superensemble models to
specific quantities of interest provides an additional calibration step
[@rykiel1996]. Individual models are typically calibrated (i.e. parameters are
estimated) based on a single response variable; however, this model might then
provide a biased or inaccurate estimate of other quantities of interest. For
example, a data-limited fisheries model might calibrate parameters by
maximizing likelihood with respect to observed and predicted catches but those
parameters may then provide a biased or inaccurate estimate of the mean and
slope of $B/B_\mathrm{MSY}$. Ensemble methods provide a final calibration step
to a quantity of interest. This ensemble calibration could further include a
loss function tailored to the goals of the model, say placing greater weight on
accuracy at low status levels than higher values.

@@

TODO

# Funding

TODO

>>>>

\noindent
Sean C. Anderson^1\*^, 
Andrew B. Cooper^1^,
Olaf P. Jensen^4^, 
C\'{o}il\'{i}n Minto^6^, 
James T. Thorson^11^,
Jessica C. Walsh^1^,
Jamie Afflerbach^2^, 
Mark Dickey-Collas^3^, 
Kristin M. Kleisner^5^, 
Catherine Longo^2^, 
Giacomo Chato Osio^7^, 
Dan Ovando^8^, 
Andrew A. Rosenberg^9^, 
Elizabeth R. Selig^10^

\noindent
\textit{order to be determined and superscripts out of order}

@@

Ecological resource managers are often faced with reconciling multiple potentially
conflicting estimates of population status and trajectory. One solution is to
average these predictions in an ensemble, but this approach ignores covariance
between models and may not optimally reduce bias and improve accuracy.
Superensemble models, commonly used for climate and weather
forecasting, may provide a superior solution. Superensembles use the predictions
from multiple models as covariates in a new statistical model. Here we evaluate
the potential for ensemble and superensemble models ('ensemble methods') to improve estimates of fish population
status and trajectory. We fit four data-limited models of population and exploitation
dynamics to marine fish stocks and combine their estimates of the mean and
slope of biomass at maximum sustainable yield ($B/B_\mathrm{MSY}$) with an
ensemble average and three superensembles: a linear model, a random forest, and
a boosted regression tree. 
We selected data-limited models for their applicability across the majority of fish stocks world wide.
We build our models on a simulated dataset of 5760
stocks and tested them with cross-validation and a global database of 
`r ram_stocks_n` stock assessments that have also been assessed with more data-intensive stock assessment models.
We find that ensemble methods substantially improve estimates of population status and trajectory.
Machine-learning superensembles performed the best for population status: accuracy improved
`r mean_sim$mach_vs_ind_mare_fold`%, correlation between predicted and true
status improved from 
`r mean_sim$ind_corr_range` to `r mean_sim$mach_corr_range`, and bias (median
proportional error) declined from 
`r mean_sim$ind_mre_range` to `r mean_sim$mach_mre_range`. 
We found similar improvements when predicting the trajectory of status. When the
simulation-trained superensembles were applied to catch data for global fish stocks, the best
superensembles improved all dimensions of performance over the next-best
individual model. Ensemble methods have the potential to
improve estimates of population status and trajectory across many taxa beyond
fish. However, they must be carefully tested, formed from a diverse set of
accurate models, and built on a dataset sufficiently representative of the
populations to which they are applied.

@@

Population status and trajectory are two of the most fundamental values to quantify in ecological resource management [e.g. @mace2008; @iucn2015]. These properties are key to everything from global assessments of extinction risk [e.g. @stuart2004; @schipper2008] to assessing hunting and fishing exploitation levels [e.g. @worm2009; @juan-jorda2011]. However, ecological resource managers are often faced with reconciling multiple uncertain and potentially conflicting estimates of status and trajectory [e.g. @brodziak2010; @branch2011]. Some models may suggest a population is at risk and declining in abundance while others may suggest it is not at risk and stable.

@@

In fisheries science, the gold standard for determining status and trajectory
of exploited fish populations are stock assessments that
incorporate all available data (e.g. catches, size and age distributions,
surveys, and tagging information) to quantify values such as 
the biomass that can produce maximum sustainable yield ($B/B_\mathrm{MSY}$).
However, these kinds of data are not available for the majority of fish
populations
[@fao2014]. Therefore, a number of models have been proposed to assess
$B/B_\mathrm{MSY}$ based on the limited data available for the majority of
fish stocks: catches and a basic understanding of population productivity [e.g.
@vasconcellos2005; @martell2013]. Recently, @rosenberg2014 investigated the
performance of four data-limited models through a large-scale simulation
experiment. The four models frequently disagreed about population status and
any one model had relatively poor predictive ability on average (e.g. Fig.
\ref{motivate}).

@@

* *mPRM* (modified panel regression model) is a modified version of the
  panel-regression model from @costello2012. Unlike the other
  models, mPRM is empirical and not mechanistic---it uses the RAM Legacy Stock
  Assessment database to fit a regression model to a series of characteristics
  of the catch time series and stock with stock-assessed $B/B_\mathrm{MSY}$ as
  the response. The model used in this paper is modified from the original in
  that it condenses the life-history categories into three categories to match
  the simulated dataset, removes the maximum catch predictor since the absolute
  catch in the simulated dataset is arbitrary, and does not implement the bias
  correction needed in @costello2012 for deriving aggregate
  estimates of median status across multiple stocks.

@@

Predictive performance can be evaluated with metrics that represent a variety
of modelling goals. For continuous response variables such as the mean and
slope of population status, performance metrics often measure some form of
bias, precision, accuracy (a combination of bias and precision), or the ability
to correctly rank or correlate across populations [e.g. @walther2005]. Here,
we measure proportional error, defined as
$(\hat{b} - b)/b$, where $\hat{b}$ and $b$ represent estimated and 'true'
(or stock-assessed) mean or slope of $B/B_\mathrm{MSY}$. We calculated median proportional
error to measure bias, median absolute proportional error to measure accuracy,
and Spearman's rank-order correlation to measure the ability to correctly rank
populations. Code to reproduce our analysis is available at
<https://github.com/datalimited/ensembles>.

@@

The superensemble models were able to improve the predictions of status by
exploiting the best properties of individual models, the covariance between
individual models, and interactions with other covariates. For example, SSCOM
had strong predictive ability when it predicted low $B/B_\mathrm{MSY}$ (Fig.
\ref{hexagon}c, Fig. \ref{partial-sim}c) and CMSY predictions were
approximately linearly related to $B/B_\mathrm{MSY}$ within the low and high prediction
clusters (Fig. \ref{partial-sim}). Superensembles also exploited the covariance
between individual model predictions. For instance, both the linear model and
GBM ensemble suggest that if mPRM and SSCOM predict high status, the true
status also tends to be high (Figs \ref{lm-coefs}, \ref{partial-2d-sim}l). The
addition of spectral density covariates helped the superensemble models correctly
predict higher status values (Fig. \ref{hexagon}g, h) by exploiting
interactions between the spectral density predictors and the individual model
predictions (Fig. \ref{partial-2d-sim}). Although the additional covariates
improved ensemble fit, performance of the ensembles was only marginally
degraded by removing these covariates (Fig. \ref{hexagon} vs. Fig.
\ref{hexagon-sim-basic}).

@@

A strength of superensembles is that they can be tailored to predict specific response
variables. For example, we built separate superensemble models of mean
$B/B_\mathrm{MSY}$ and the slope of $B/B_\mathrm{MSY}$. The same set of model
weights or non-linear relationships need not hold across different response
variables. For example, SSCOM contributed little to the GBM superensemble estimate
of status at higher levels of predicted $B/B_\mathrm{MSY}$, but contributed
strongly to estimates of slope (Fig. \ref{fig:partial-sim-slope}). Formally,
fitting superensemble models to
specific quantities of interest provides an additional calibration step
[@rykiel1996]. Individual models are typically calibrated (i.e. parameters are
estimated) based on a single response variable; however, this model might then
provide a biased or inaccurate estimate of other quantities of interest. For
example, a data-limited fisheries model might calibrate parameters by
maximizing likelihood with respect to observed and predicted catches but those
parameters may then provide a biased or inaccurate estimate of the mean and
slope of $B/B_\mathrm{MSY}$. Ensemble methods provide a final calibration step
to a quantity of interest. This ensemble calibration could further include a
loss function tailored to the goals of the model, say placing greater weight on
accuracy at low status levels than higher values.

@@

We thank members of Phase I of the working group 'INSERT NAME HERE' who
contributed to developing the data-limited methods and simulations used in our
analysis.
We thank J.A. Hutchings for helpful discussions on ensemble methods for
assessing population status.
We thank the Gordon and Betty Moore Foundation for funding of the working group
'INSERT NAME HERE'.
